Machine Learning for Predictive Auto-Tuning

### 2011-12-13 paper/draft1.tex
Edits and comments
minor intro edit

In this rapidly evolving landscape, the construction of \emph{general}
development tools and libraries that \emph{fully utilize} system resources
remains a daunting task \cite{?}. Even within specialized architectures from
the \emph{same} vendor, like NVIDIA's Graphics Processing Units (GPUs) and the
Compute Unified Device Architecture (CUDA) \cite{CUDA}, many developers default
to massive amounts of manual labor to optimize CUDA code to specific input
domains. In addition, hand-tuning rarely generalizes well to new hardware
generations or different input domains, and it can also be error-prone or far
from optimal. One of the reason is that kernels can produce staggeringly large
optimization spaces \cite{K. Datta et al. Stencil computation optimization and
auto-tuning on state-of-the-art multicore architectures}. The problem is
further compounded by the fact that these spaces can be highly discontinuous
\cite{RyooHwuCGO2008}, difficult to explore and holding quasi-optimal solutions
at the edge of ''performance cliffs'' (in part due to hard resource
restrictions, like the register file size).

>>>>

In this rapidly evolving landscape, the construction of \emph{general}
development tools and libraries that \emph{fully utilize} system resources
remains a daunting task \cite{?}. Even within specialized architectures from
the \emph{same} vendor, like NVIDIA's Graphics Processing Units (GPUs) and the
Compute Unified Device Architecture (CUDA) \cite{CUDA}, many developers default
to massive amounts of manual labor to optimize CUDA code to specific input
domains. In addition, hand-tuning rarely generalizes well to new hardware
generations or different input domains, and it can also be error-prone or far
from optimal. Kernels can produce staggeringly large
optimization spaces \cite{K. Datta et al. Stencil computation optimization and
auto-tuning on state-of-the-art multicore architectures}, in which performance can vary
widely and discontinously~\citep{RyooHwuCGO2008}.
The best kernels are often at the edge of ``performance cliffs'' induced by
device-specific hardware limits (e.g. shared memory size, register file size, block size).
